MSnbase2 benchmarking

### 2017-01-20 docs/articles/MSnbase-demo.Rnw

\input{NoteAboutSpeedAndMemory.tex}

@@

\begin{figure}[htb]
    \includegraphics[width=\linewidth]{./plot2d-figure.png}
\caption{Illustration of the \Rfunction{plot2d} output. }
\label{fig:plot2d}
\end{figure}

@@

\begin{figure}[htb]
    \includegraphics[width=\linewidth]{./plotDensity-figure.png}
\caption{ Illustration of the \Rfunction{plotDensity} output. }
\label{fig:plotDensity}
\end{figure}

@@

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=.7\linewidth]{./plotMzDelta-pride12011.pdf}
    \caption{Illustration of the \Rfunction{plotMzDelta} output for
      the PRIDE experiment 12011, as in figure 4A from
      \cite{Foster11}. }
\label{fig:plotMzDelta-pride12011}
  \end{center}
\end{figure}

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=.6\linewidth]{./tmt10comp.pdf}
\caption{TMT 10-plex quantitation. }
\label{fig:tmt10}
\end{figure}

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=4cm]{./itraqchem.pdf}
    \caption{ iTRAQ 4-plex isobaric tags reagent consist of three
      parts: (1) a charged reporter group (MZ of 114, 115, 116 and
      117) that is unique to each of the four reagents (red), (2) an
      uncharged mass balance group (28-31 Da) (blue)and (3) a peptide
      reactive group (NHS ester) that binds to the peptide.  In case
      of incomplete dissociation, the reporter and balance groups
      produce a specific peaks at MZ 145.  }
\label{fig:itraqchem}
  \end{center}
\end{figure}

>>>>

\subsection{Speed and memory requirements}

Raw mass spectrometry file are generally several hundreds of MB large
and most of this is used for binary raw spectrum data. As such, data
containers can easily grow very large and thus require large amounts
of RAM.  This requirement is being tackled by avoiding to load the raw
data into memory and using on-disk random access to the content of
\texttt{mzXML}/\texttt{mzML} data files on demand.  When focusing on
reporter ion quantitation, a direct solution for this is to trim the
spectra using the \Robject{trimMz} method to select the area of
interest and thus substantially reduce the size of the
\Robject{Spectrum} objects. This is illustrated in section
\ref{sec:trim} on page \pageref{trimMz-example}.

The independent handling of spectra is ideally suited for parallel
processing.  The \Rfunction{quantify} method now performs reporter
peaks quantitation in parallel.  More functions are being updated.

Finally, recent developmenets in version 2 of the package have solved
the memory issue by implementing and \textit{on-disk} version the of
data class storing raw data (\Rclass{MSnExp}, see
section~\ref{sec:msnexp}), where the spectra a accessed on-disk only
when required. The \textit{benchmarking} vignette compares the on-disk
and in-memory implemenatations.

@@

\begin{figure}[htb]
    \includegraphics[width=\linewidth]{./Figures/plot2d-figure.png}
\caption{Illustration of the \Rfunction{plot2d} output. }
\label{fig:plot2d}
\end{figure}

@@

\begin{figure}[htb]
    \includegraphics[width=\linewidth]{./Figures/plotDensity-figure.png}
\caption{ Illustration of the \Rfunction{plotDensity} output. }
\label{fig:plotDensity}
\end{figure}

@@

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=.7\linewidth]{./Figures/plotMzDelta-pride12011.pdf}
    \caption{Illustration of the \Rfunction{plotMzDelta} output for
      the PRIDE experiment 12011, as in figure 4A from
      \cite{Foster11}. }
\label{fig:plotMzDelta-pride12011}
  \end{center}
\end{figure}

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=.6\linewidth]{./Figures/tmt10comp.pdf}
\caption{TMT 10-plex quantitation. }
\label{fig:tmt10}
\end{figure}

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=4cm]{./Figures/itraqchem.pdf}
    \caption{ iTRAQ 4-plex isobaric tags reagent consist of three
      parts: (1) a charged reporter group (MZ of 114, 115, 116 and
      117) that is unique to each of the four reagents (red), (2) an
      uncharged mass balance group (28-31 Da) (blue)and (3) a peptide
      reactive group (NHS ester) that binds to the peptide.  In case
      of incomplete dissociation, the reporter and balance groups
      produce a specific peaks at MZ 145.  }
\label{fig:itraqchem}
  \end{center}
\end{figure}
